Type Refinements for the Working Class

### 2016-01-01 paper.tex
Add possibility modality; fix rule for functorial lifting of modalities (#13)

  \begin{gather*}
    \infer[\textit{extract}]{
      \RMemEq{\Upsilon}{M}{N}{\phi}
    }{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
    }
    \qquad
    \infer=[\textit{duplicate}]{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\Indep{\phi}{a}}{a}}
    }{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
    }
    \\[6pt]
    \infer[\textit{lift}]{
      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
      \implies
      \RMemEq{\Upsilon}{M}{N}{\Indep{\psi}{a}}
    }{
      \forall\Member{M,N}{\Exprs_\tau(\Upsilon)}.\
      \RMemEq{\Upsilon}{M}{N}{\phi}
      \implies
      \RMemEq{\Upsilon}{M}{N}{\psi}
    }
  \end{gather*}
\end{thm}
\begin{proof}
  The first two inferences are obviously evident. To validate \textit{lift},
  suppose that $\RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}$; by inversion, we
  have an $\Member{L}{\Exprs_\tau(\Upsilon\setminus a)}$ such that
  $\RMemEq{\Upsilon}{L}{M}{\phi}$ and $\RMemEq{\Upsilon}{L}{N}{\phi}$.  Now, it
  suffices to exhibit an $\Member{L'}{\Exprs_\tau(\Upsilon\setminus a)}$ such
  that $\RMemEq{\Upsilon}{L'}{M}{\psi}$ and $\RMemEq{\Upsilon}{L'}{N}{\psi}$;
  but we may simply choose $\Match{L'}{L}$ and discharge the remaining subgoals
  by appealing to the premise of the rule.
\end{proof}

@@

>>>>

  \begin{gather*}
    \infer[\textit{extract}]{
      \RMemEq{\Upsilon}{M}{N}{\phi}
    }{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
    }
    \qquad
    \infer=[\textit{duplicate}]{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\Indep{\phi}{a}}{a}}
    }{
      \RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}
    }
    \\[6pt]
    \infer[\textit{lift}]{
      \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}.\
      \forall\Member{M,N}{\Exprs_\tau(\Upsilon')}.\
      \RMemEq{\Upsilon'}{M}{N}{\Indep{\phi}{a}}
      \implies
      \RMemEq{\Upsilon'}{M}{N}{\Indep{\psi}{a}}
    }{
      \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}.\
      \forall\Member{M,N}{\Exprs_\tau(\Upsilon')}.\
      \RMemEq{\Upsilon'}{M}{N}{\phi}
      \implies
      \RMemEq{\Upsilon'}{M}{N}{\psi}
    }
  \end{gather*}
\end{thm}
\begin{proof}
  The first two inferences are obviously evident. To validate \textit{lift},
  suppose that $\RMemEq{\Upsilon}{M}{N}{\Indep{\phi}{a}}$; by inversion, we
  have an $\Member{L}{\Exprs_\tau(\Upsilon\setminus a)}$ such that
  $\RMemEq{\Upsilon}{L}{M}{\phi}$ and $\RMemEq{\Upsilon}{L}{N}{\phi}$.  Now, it
  suffices to exhibit an $\Member{L'}{\Exprs_\tau(\Upsilon\setminus a)}$ such
  that $\RMemEq{\Upsilon}{L'}{M}{\psi}$ and $\RMemEq{\Upsilon}{L'}{N}{\psi}$;
  but we may simply choose $\Match{L'}{L}$ and discharge the remaining subgoals
  by appealing to the premise of the rule.
\end{proof}

@@

\subsubsection{Nominal Possibility}

\emph{Nominal possibility} expresses the idea that there \emph{exists} an
extension of the current symbol context at which a refinement is inhabited.
However, it is not possible to define a useful possibility modality in the same
way as the others, without imposing requirements on the abt signature. In
particular, we need to introduce a new sort and a new canonical operator for
introducing possibility:
\[
  \infer{
    \IsSort{\Diamond\tau}
  }{
    \IsSort{\tau}
  }\qquad
  \infer{
    \IsCanOperator{\Upsilon}{\Diamond[\vec{\sigma}]}{
      \MkArity{
        \MkValence{\vec{\sigma}}{}{\tau}
      }{\Diamond\tau}
    }
  }{
  }
\]

Then, we can explain the following refinement formation rule:
\[
  \infer={
    \Refines{\Upsilon}{\Diamond\phi}{\Diamond\tau}
  }{
    \Refines{\Upsilon}{\phi}{\tau}
  }
\]

We will begin by defining a refinement $\Diamond_0\phi$ on values, for any renaming
\begin{tikzcd}[cramped]
  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd}%
:
\[
  \infer={
    \RMemEq{\Upsilon'}{
      \App{
        \Diamond[\vec{\sigma}]
      }{
        \MkAbs{\vec{u}}{}{M}
      }
    }{
      \App{
        \Diamond[\vec{\sigma}]
      }{
        \MkAbs{\vec{v}}{}{N}
      }
    }{\Diamond_0\phi}
  }{
    \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon'}\cup\vec{u}\cup\vec{v}}.\
    \RMemEq{\Upsilon',\vec{w}:\vec{\sigma}}{\Rename{(\vec{u}\mapsto\vec{w}}{M})}{\Rename{(\vec{v}\mapsto\vec{w})}{N}}{\phi}
  }
\]

Then, the possibility modality is defined as the computational extension of
$\Diamond_0-$, $\Define{\Diamond\phi}{\VToCType\left(\Diamond_0\phi\right)}$.

\begin{thm}

  Nominal possibility is an idempotent monadic modality, in the sense that the
  following rules are valid:
%
  \[
    \infer[\textit{return}]{
      \RMemEq{\Upsilon}{
        \App{
          \Diamond[\cdot]
        }{
          \MkAbs{}{}{M}
        }
      }{
        \App{
          \Diamond[\cdot]
        }{
          \MkAbs{}{}{N}
        }
      }{\Diamond\phi}
    }{
      \RMemEq{\Upsilon}{M}{N}{\phi}
    }
  \]
  \[
    \infer=[\textit{join}]{
      \RMemEq{\Upsilon}{
        \App{
          \Diamond[\vec{\sigma}_0,\vec{\sigma}_1]
        }{
          \MkAbs{\vec{u}_0,\vec{u}_1}{}{M}
        }
      }{
        \App{
          \Diamond[\vec{\sigma}_0,\vec{\sigma}_1]
        }{
          \MkAbs{\vec{v}_0,\vec{v}_1}{}{M}
        }
      }{\Diamond\phi}
    }{
      \RMemEq{\Upsilon}{
        \App{
          \Diamond[\vec{\sigma}_0]
        }{
          \MkAbs{\vec{u}_0}{}{
            \App{
              \Diamond[\vec{\sigma}_1]
            }{
              \MkAbs{\vec{u}_1}{}{M}
            }
          }
        }
      }{
        \App{
          \Diamond[\vec{\sigma_0}]
        }{
          \MkAbs{\vec{v}_0}{}{
            \App{
              \Diamond[\vec{\sigma}_1]
            }{
              \MkAbs{\vec{v}_1}{}{N}
            }
          }
        }
      }{\Diamond\Diamond\phi}
    }
  \]
  \[
    \infer[\textit{lift}]{
      \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}.\
      \forall\Member{M,N}{\Exprs_{\Diamond\tau}(\Upsilon')}.\
      \RMemEq{\Upsilon'}{M}{N}{\Diamond\phi}
      \implies
      \RMemEq{\Upsilon'}{M}{N}{\Diamond\psi}
    }{
      \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}.\
      \forall\Member{M,N}{\Exprs_\tau(\Upsilon')}.\
      \RMemEq{\Upsilon'}{M}{N}{\phi}
      \implies
      \RMemEq{\Upsilon'}{M}{N}{\psi}
    }
  \]
\end{thm}
\begin{proof}

  As in Theorem~\ref{thm:independence-comonadic}, the validity first two rules
  is obvious. To demonstrate the \textit{lift} rule, fix a renaming
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &
    \IMode{\Upsilon'}\arrow[r,hook,"\IMode{\rho'}"] &
    \IMode{\Upsilon''}\equiv\OMode{\Upsilon', \vec{w}:\vec{\sigma}}
  \end{tikzcd}
  and $\Member{M,N}{\Exprs_\tau(\Upsilon'')}$ such that
  $\RMemEq{\Upsilon''}{M}{N}{\phi}$; then it suffices to show that
  $\RMemEq{\Upsilon'}{\App{\Diamond[\vec{\sigma}]}{\MkAbs{\vec{w}}{}{M}}}{\App{\Diamond[\vec{\sigma}]}{\MkAbs{\vec{w}}{}{N}}}{\Diamond\psi}$, which is to say,
  $\RMemEq{\Upsilon''}{M}{N}{\psi}$. But this follows directly from the premise, setting $\Match{\Upsilon'}{\Upsilon''}$.
